---
title: "final_project"
author: "Kangkang Zhang"
date: "12/7/2018"
output: github_document
---

```{r, message = FALSE}
library(tidyverse)
```

Import data and check missing values.

```{r, message = FALSE}
cancer_data_raw = read_csv("./data/Cancer_Registry.csv")

#check missing data
cancer_data_raw %>%
    janitor::clean_names() %>% 
  summarise_all(funs(sum(is.na(.)))) %>% 
  gather(term, na_num, avg_ann_count:birth_rate) %>% 
  filter(na_num > 0) %>% 
  mutate(percent = na_num/sum(na_num))
```

Exclude variable PctSomeCol18_24 since its missing rate is 75%. Fill in missing values in pct_employed16_over and pct_private_coverage_alone using mean values of all none-NA obs.

```{r}
cancer_data = cancer_data_raw %>% 
  janitor::clean_names() %>%
  select(target_death_rate, everything(), - pct_some_col18_24) %>%
  mutate(pct_employed16_over = ifelse(is.na(pct_employed16_over), 
                                      mean(pct_employed16_over, na.rm = TRUE), pct_employed16_over),
         pct_private_coverage_alone = ifelse(is.na(pct_private_coverage_alone), 
                                      mean(pct_private_coverage_alone, na.rm = TRUE), pct_private_coverage_alone))

```

extract names of numeric variable

```{r}
vars = cancer_data %>% 
  keep(is.numeric) %>% 
  names()
```

summary statistics for each numeric variable

```{r}
cancer_data %>%
  select(vars) %>% 
  summary() %>% 
  broom::tidy() %>%
  select(- Var1) %>% 
  separate(n, c("stat", "value"), sep = ":") %>% 
  spread(key = Var2, value = value) %>% 
  knitr::kable()
```

check normality of each variable

```{r}
#variable 1:9
cancer_data %>%
  select(vars[1:9]) %>% 
  gather() %>%   
  ggplot(aes(value)) +                    
    facet_wrap(~ key, scales = "free") +  
    geom_density(na.rm = TRUE) 
```

```{r}
#variable 10:18
cancer_data %>%
  select(vars[10:18]) %>% 
  gather() %>%   
  ggplot(aes(value)) +                    
    facet_wrap(~ key, scales = "free") +  
    geom_density(na.rm = TRUE) 
```

```{r}
#variable 19:27
cancer_data %>%
  select(vars[19:27]) %>% 
  gather() %>%   
  ggplot(aes(value)) +                    
    facet_wrap(~ key, scales = "free") +  
    geom_density(na.rm = TRUE) 
```

```{r}
#variable 28:31
cancer_data %>%
  select(vars[28:31]) %>% 
  gather() %>%   
  ggplot(aes(value)) +                    
    facet_wrap(~ key, scales = "free") +  
    geom_density(na.rm = TRUE) 
```

check linear relationship

```{r}
#variable 2:10
cancer_data %>%
  select(vars[1:10]) %>% 
  gather(key = key, value, vars[2:10]) %>%   
  ggplot(aes(x = value, y = target_death_rate)) +                    
    facet_wrap(~ key, scales = "free") +  
    geom_point(na.rm = TRUE) +
    geom_smooth(se = FALSE, method = "lm") 
```

variable median_age has some abnormal value, 30 obs are large than 150. Exclude those obs?

```{r}
cancer_data %>%
  select(median_age) %>% 
  filter(median_age > 150)  #result is the same as larger than 100

cancer_data_df = cancer_data %>%
  filter(median_age <= 150)
  
```

check the linear relationship between median_age and responce.

```{r}
cancer_data %>%
  select(target_death_rate, median_age) %>%
  filter(median_age <= 150) %>% 
  ggplot(aes(x = median_age, y = target_death_rate)) +                    
  geom_point(na.rm = TRUE) +
  geom_smooth(se = FALSE, method = "lm") 
```


```{r}
#variable 11:19
cancer_data %>%
  select(target_death_rate, vars[11:19]) %>% 
  gather(key = key, value, vars[11:19]) %>%   
  ggplot(aes(x = value, y = target_death_rate)) +                    
    facet_wrap(~ key, scales = "free") +  
    geom_point(na.rm = TRUE) +
  geom_smooth(se = FALSE, method = "lm") 
```

```{r}
#variable 20:28
cancer_data %>%
  select(target_death_rate, vars[20:28]) %>% 
  gather(key = key, value, vars[20:28]) %>%   
  ggplot(aes(x = value, y = target_death_rate)) +                    
    facet_wrap(~ key, scales = "free") +  
    geom_point(na.rm = TRUE) +
  geom_smooth(se = FALSE, method = "lm") 
```

```{r}
#variable 29:31
cancer_data %>%
  select(target_death_rate, vars[29:31]) %>% 
  gather(key = key, value, vars[29:31]) %>%   
  ggplot(aes(x = value, y = target_death_rate)) +                    
    facet_wrap(~ key, scales = "free") +  
    geom_point(na.rm = TRUE) +
  geom_smooth(se = FALSE, method = "lm") 
```

check correlation

```{r}
cancer_data_df %>%
  keep(is.numeric) %>% 
  cor() %>% 
  knitr::kable()

#Descending order according to correlocation with target variable
cancer_data_df %>%
  keep(is.numeric) %>% 
  cor() %>% 
  as_tibble() %>% 
  knitr::kable()

```

exclude variable pct_private_coverage_alone

```{r}
cancer_data_df = cancer_data_df %>%
  keep(is.numeric) %>% 
  dplyr::select(-pct_private_coverage_alone)
```


Use lasso to select varibles in all variables

```{r}
library('faraway')
library('MASS')                           # for lm.ridge()
library('glmnet')                         # for glmnet()
library('reshape2')

cancer_data_df1<- data.frame(cancer_data_df)

# Start with the full model
mult.fit <- lm(target_death_rate ~ ., data=cancer_data_df1)
summary(mult.fit)

Y <- cancer_data_df1[,1]

X <- as.matrix(cancer_data_df1[,-1])

set.seed(1)

#split dataset into train and test 8:2
train<-sample(1:nrow(X),nrow(X)*0.8)

test<-(-train)

Y.test<-Y[test]

#Lasso regression
grid <- 10^seq(5,-2, length=100)
lasso1<- glmnet(X,Y, alpha =1, lambda =grid)

# Cross-validation
set.seed(2)
cv.out<-cv.glmnet(X[train,],Y[train])
plot(cv.out)

best.lambda<-cv.out$lambda.1se

#lasso.pred=predict(lasso1,s=best.lambda,newx=X[test,])
#mean((lasso.pred-Y.test)^2)

# Fit a Lasso model with all observations with the best lambda
lasso2<- glmnet(X, Y, alpha =1, lambda=best.lambda)
coef(lasso2)
coef(mult.fit)

# Fraction of deviance explained
# Similar interpretation to R-squared: % variance explained by non-zero variables variables
lasso2$dev.ratio

# Compare LS, Ridge and Lasso regression coefficients
res_ls_ridge_lasso<- cbind(coef(mult.fit),coef(lasso2))
colnames(res_ls_ridge_lasso) <- c("LS", "Lasso")
res_ls_ridge_lasso

# Using the entire data, fit Lasso regressions using the lambda grid.
lasso3 <- glmnet(X,Y, alpha=1, lambda=grid)

# Save the estimated 'standardized' coefficients for all 7 predictors without the intercept that is not of interest.
coef_lasso3 <- coef(lasso3)[-1,]
# Transpose the matrix
coef_lasso3_mat <- t(as.matrix(coef_lasso3))
# Rename and sort the matrix by asceding  grid
rownames(coef_lasso3_mat) <- grid
coef_lasso3_mat_sort <- coef_lasso3_mat[order(grid),]

# Plot using different colors
matplot(coef_lasso3_mat_sort,type="l",lty=1,xlim=c(0,50),
        xlab="lambda",ylab="coefficient",col=1:7) 
### add legend
legend('bottomright', inset=.05, legend=colnames(coef_lasso3_mat_sort), 
       pch=20, cex=0.8, col=1:7)
```



Variable selection after stepwise and AIC

```{r}
cancer_data_df2 = cancer_data_raw %>% 
  dplyr::select(TARGET_deathRate,incidenceRate,medIncome, MedianAgeFemale,
    PctNoHS18_24, PctHS18_24, PctHS25_Over, PctBachDeg25_Over,
    PctUnemployed16_Over, PctPrivateCoverage, PctEmpPrivCoverage, 
    PctWhite, PctOtherRace) %>%
  janitor::clean_names() %>% 
  data.frame()

mult.fit <- lm(target_death_rate ~ ., data=cancer_data_df2)
summary(mult.fit)

Y <- cancer_data_df2[,1]

X <- as.matrix(cancer_data_df2[,-1])

set.seed(1)

train<-sample(1:nrow(X),nrow(X*0.8))

test<-(-train)

Y.test<-Y[test]


lasso1<- glmnet(X[train ,],Y[train], alpha =1, lambda =grid)


# Cross-validation
cv.out<-cv.glmnet(X[train,],Y[train])
plot(cv.out)

best.lambda<-cv.out$lambda.1se

#lasso.pred=predict(lasso1,s=best.lambda,newx=X[test,])
#mean((lasso.pred-Y.test)^2)

# Fit a Lasso model with all observations with the best lambda
lasso2<- glmnet(X, Y, alpha =1, lambda=best.lambda)
coef(lasso2)
coef(mult.fit)

# Fraction of deviance explained
# Similar interpretation to R-squared: % variance explained by non-zero variables variables
lasso2$dev.ratio

# Compare LS, Ridge and Lasso regression coefficients
res_ls_ridge_lasso<- cbind(coef(mult.fit),coef(lasso2))
colnames(res_ls_ridge_lasso) <- c("LS", "Lasso")
res_ls_ridge_lasso


lasso3 <- glmnet(X,Y, alpha=1, lambda=grid)

# Save the estimated 'standardized' coefficients for all 7 predictors without the intercept that is not of interest.
coef_lasso3 <- coef(lasso3)[-1,]
# Transpose the matrix
coef_lasso3_mat <- t(as.matrix(coef_lasso3))
# Rename and sort the matrix by asceding  grid
rownames(coef_lasso3_mat) <- grid
coef_lasso3_mat_sort <- coef_lasso3_mat[order(grid),]

# Plot using different colors
matplot(coef_lasso3_mat_sort,type="l",lty=1,xlim=c(0,50),
        xlab="lambda",ylab="coefficient",col=1:7) 
### add legend
legend('bottomright', inset=.05, legend=colnames(coef_lasso3_mat_sort), 
       pch=20, cex=0.8, col=1:7)

```

The lm model is :
```{r}
cancer_data_df %>% 
  dplyr::select(target_death_rate, incidence_rate,med_income, pct_hs18_24, pct_hs25_over,
                pct_bach_deg25_over,
                pct_unemployed16_over, pct_private_coverage, pct_white, pct_other_race) %>% 
lm(target_death_rate~., data = .) %>% 
  summary()
```



selected 15 predictors
```{r}
cancer_data_dff = cancer_data_df %>% 
  dplyr::select(target_death_rate, incidence_rate, study_per_cap, med_income, poverty_percent,
         median_age, pct_bach_deg25_over, pct_unemployed16_over, pct_emp_priv_coverage, pct_public_coverage, pct_white, pct_black,
         pct_asian)
```


selected 11 predictors(exclude income and 3 education related vars)
```{r}
cancer_data_1df = cancer_data %>% 
  dplyr::select(target_death_rate, incidence_rate, study_per_cap, poverty_percent,
         median_age, pct_bach_deg25_over, pct_unemployed16_over, pct_emp_priv_coverage, pct_public_coverage, pct_white, pct_black,
         pct_asian)

mult.fit <- lm(target_death_rate ~ ., data = cancer_data_1df)
step(mult.fit, direction = 'both')


# Leaps function provides all-subsets analysis
cancer_data_dfff1 = cancer_data_dff1 %>% 
  dplyr::select(target_death_rate, incidence_rate, poverty_percent,
         pct_bach_deg25_over, pct_unemployed16_over, pct_emp_priv_coverage, pct_public_coverage, pct_white) %>% 
  as.data.frame()

# Printing the 2 best models of each size, using the Cp criterion:
leaps(x = cancer_data_dfff1[,2:8], y = cancer_data_dfff1[,1], nbest=2, method="Cp")


# Printing the 2 best models of each size, using the adjusted R^2 criterion:
leaps(x = cancer_data_dfff1[,2:8], y = cancer_data_dfff1[,1], nbest=2, method="adjr2")

# Summary of models for each size (one model per size)
b<-regsubsets(target_death_rate ~ ., data = cancer_data_dfff1)
(rs<-summary(b))

# Plots of Cp and Adj-R2 as functions of parameters
par(mar=c(4,4,1,1))
par(mfrow=c(1,2))

plot(2:8, rs$cp, xlab="No. of parameters", ylab="Cp Statistic")
abline(0,1)

plot(2:8, rs$adjr2, xlab="No of parameters", ylab="Adj R2")
```


```{r}
# rstandard function gives the INTERNALLY studentized residuals 
mult.fit1 = lm(target_death_rate~incidence_rate+poverty_percent+pct_bach_deg25_over+pct_unemployed16_over+               pct_emp_priv_coverage, data = cancer_data_1df)


stu_res<-rstandard(mult.fit1)
outliers_y<-stu_res[abs(stu_res)>2.5]

# Measures of influence:
# Gives DFFITS, Cook's Distance, Hat diagonal elements, and others.

influence.measures(mult.fit1)

# Look at the Cook's distance lines and notice obs 5 and 28 as potential Y outliesrs / influential points

par(mfrow=c(2,2))
plot(mult.fit1)

# Examine results with and without observations 5 and 28 that have very high survivals (>2000)
summary(mult.fit1)


# Remove observations 5 and 28
cancer_df_remove<-cancer_data_1df[c(-1221,-1366, -282),]

mult.fit_remove<- lm(target_death_rate~incidence_rate+poverty_percent+pct_bach_deg25_over+pct_unemployed16_over+               pct_emp_priv_coverage, data = cancer_df_remove) 

summary(mult.fit_remove)

influence.measures(mult.fit_remove)

par(mfrow=c(2,2))
plot(mult.fit_remove)
```

```{r}
cancer_data_1df %>% 
  mutate(log_poverty = sqrt(poverty_percent)) %>% 
  ggplot(aes(x = log_poverty))+
  geom_density()
```

```{r}
library(MASS)
fit1 <- lm(Survival ~ Bloodclot, data=data_surg)
boxcox(fit1)  # default grid of lambdas is -2 to 2 by 0.1

# Could change grid of lambda values
boxcox(fit1, lambda = seq(-3, 3, by=0.25) ) 


# Box Cox for multiple regression
mult.fit1 <- lm(Survival ~ Bloodclot + Progindex + Enzyme + Liver + Age + Gender + Alcmod + Alcheav, data=data_surg) 
summary(mult.fit1)

boxcox(mult.fit1) 
plot(mult.fit1)


mult.fit2 <- lm(Lnsurvival ~ Bloodclot + Progindex + Enzyme + Liver + Age + Gender + Alcmod + Alcheav, data=data_surg) 
summary(mult.fit2)
boxcox(mult.fit2) 
plot(mult.fit2)

```

