---
title: "Final Project R code"
output: html_document
---

```{r, message = FALSE}
library(tidyverse)
library(HH)
library(leaps)
library(boot)
library(broom)
library(ggpubr)
library(reshape2)
```

#Import and Tidy data

Import data and check missing values.

```{r, message = FALSE}
cancer_data_raw = read_csv("./data/Cancer_Registry.csv")

#check missing data
cancer_data_raw %>%
    janitor::clean_names() %>% 
  summarise_all(funs(sum(is.na(.)))) %>% 
  gather(term, na_num, avg_ann_count:birth_rate) %>% 
  filter(na_num > 0) %>% 
  mutate(percent = na_num/sum(na_num))
```

Exclude variable PctSomeCol18_24 since its missing rate is 75%. Fill in missing values in pct_employed16_over and pct_private_coverage_alone using mean values of all none-NA obs.

```{r}
cancer_data = cancer_data_raw %>% 
  janitor::clean_names() %>%
  dplyr::select(target_death_rate, everything(), - pct_some_col18_24) %>%
  mutate(pct_employed16_over = ifelse(is.na(pct_employed16_over), 
                                      mean(pct_employed16_over, na.rm = TRUE), pct_employed16_over),
         pct_private_coverage_alone = ifelse(is.na(pct_private_coverage_alone), 
                                      mean(pct_private_coverage_alone, na.rm = TRUE), pct_private_coverage_alone))

```

Select 11 predictors of interest(with no missing data).

```{r}
cancer_data_df= cancer_data %>% 
  dplyr::select(target_death_rate, incidence_rate, poverty_percent,
         median_age_female, pct_bach_deg25_over, pct_unemployed16_over, pct_emp_priv_coverage, pct_public_coverage, pct_white, pct_black,
         pct_asian, pct_other_race)

#Fit a regression using all predictors
mult.fit <- lm(target_death_rate ~ ., data = cancer_data_df)
summary(mult.fit)
```

#Descriptive statistics

```{r}
#density plot of target
cancer_data_df %>%
  dplyr::select(target_death_rate) %>% 
  ggplot(aes(x = target_death_rate)) +                    
    geom_density(na.rm = TRUE) +
  theme_bw()

#boxplot of predictors
cancer_data_df %>%
  dplyr::select(-target_death_rate) %>%  
  gather() %>%   
  ggplot(aes(x = value, y = value)) +                    
    facet_wrap(~ key, scales = "free") +  
    geom_boxplot(na.rm = TRUE) +
    theme_bw()
```

```{r}
#descriptive statistics
descriptive = function(variable = as.character(), data = cancer_data_df){
    df = summary(data[[variable]]) %>% tidy()
    df$sd = sd(data[[variable]])
    df[c(4, 7, 1, 2, 3, 5, 6)]
}

a = tibble(
    variable = names(cancer_data_df),
    desc = map(variable, ~descriptive(variable = .x))
) %>% 
    unnest()
```

```{r}
#correlation heat map
data_pct = cancer_data_df %>%
  dplyr::select(- target_death_rate)

cormat = round(cor(data_pct), 2)
melted_cormat = melt(cormat)

ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "steelblue", high = "lightcoral", mid = "white", 
  midpoint = 0, limit = c(-1,1), space = "Lab", 
  name="Pearson\nCorrelation") +
  geom_text(aes(Var2, Var1, label = value), color = "black", size = 3) +
  theme(axis.text.x = element_text(angle = 20, vjust = 1, 
    size = 8, hjust = 1)) +
  labs(
    x = "",
    y = "")
```

#Model Building  

##Variable Selection

###Backward Elimination

```{r}
# No pct_asian
step1<-update(mult.fit, . ~ . -pct_asian)
summary(step1)

# No pct_black
step2<-update(step1, . ~ . -pct_black)
summary(step2)

#all predictors have p-value less than 0.05, procedure stops
reg1 = lm(formula = target_death_rate ~ incidence_rate + poverty_percent + 
    median_age_female + pct_bach_deg25_over + pct_unemployed16_over + 
    pct_emp_priv_coverage + pct_public_coverage + pct_white + 
    pct_other_race, data = cancer_data_df)

summary(reg1)
vif(reg1) #vif < 10
```

###AIC based procedures 

```{r}
mult.fit <- lm(target_death_rate ~ ., data = cancer_data_df)
step(mult.fit, direction = 'both')

#the model is
reg2 = lm(formula = target_death_rate ~ incidence_rate + poverty_percent + 
    median_age_female + pct_bach_deg25_over + pct_unemployed16_over + 
    pct_emp_priv_coverage + pct_public_coverage + pct_white + 
    pct_other_race, data = cancer_data_df)

summary(reg2)
vif(reg2)

#reg2 is the same as reg1
```

###All-subset Analysis

```{r}
cancer_data_df1 = cancer_data_df %>% 
  as.data.frame()

# Summary of models for each size (one model per size)
b<-regsubsets(target_death_rate ~ ., data = cancer_data_df1, nvmax = 11)
(rs<-summary(b))

# Plots of Cp and Adj-R2 as functions of parameters
par(mar=c(4,4,1,1))
par(mfrow=c(1,2))

plot(2:12, rs$cp, xlab="No. of parameters", ylab="Cp Statistic")
abline(0,1)

plot(2:12, rs$adjr2, xlab="No. of parameters", ylab="Adj R Square")

rbind(rs$cp, rs$adjr2)
```


```{r}
#AIC 
AIC(mult.fit)
stp1 = update(mult.fit, . ~ . -pct_asian)
AIC(stp1)
stp2 = update(stp1, . ~ . -pct_black)
AIC(stp2)
stp3 = update(stp2, . ~ . -pct_emp_priv_coverage)
AIC(stp3)
stp4 = update(stp3, . ~ . -pct_public_coverage)
AIC(stp4)
stp5 = update(stp4, . ~ . -median_age_female)
AIC(stp5)
stp6 = update(stp5, . ~ . -pct_white)
AIC(stp6)
stp7 = update(stp6, . ~ . -pct_unemployed16_over)
AIC(stp7)
stp8 = update(stp7, . ~ . -pct_other_race)
AIC(stp8)
stp9 = update(stp8, . ~ . -poverty_percent)
AIC(stp9)
stp10 = update(stp9, . ~ . -incidence_rate)
AIC(stp10)

#BIC
k = log(length(cancer_data_df$target_death_rate))

AIC(mult.fit, k = k)
AIC(stp1, k = k)
AIC(stp2, k = k)
AIC(stp3, k = k)
AIC(stp4, k = k)
AIC(stp5, k = k)
AIC(stp6, k = k)
AIC(stp7, k = k)
AIC(stp8, k = k)
AIC(stp9, k = k)
AIC(stp10, k = k)

#the final model is
final_model = lm(formula = target_death_rate ~ incidence_rate + poverty_percent + 
    pct_bach_deg25_over + pct_unemployed16_over + pct_white + 
    pct_other_race, data = cancer_data_df)

vif(final_model)

confint(final_model, level=0.95)

```

Model with 7 parameters has the smallest BIC, and its Adjusted R-Sq is comparatively larger. We decide to choose it as our final model.

##Model Diagnostics

```{r}
#box-cox
boxcox(final_model)

# rstandard function gives the INTERNALLY studentized residuals 
stu_res<-rstandard(final_model)
outliers_y<-stu_res[abs(stu_res)>2.5]
outliers_y

influence.measures(final_model)

par(mfrow=c(2,2))
plot(final_model)

# Remove obs 1221, 1366, 282
cancer_df_remove<-cancer_data_df[c(-1221, -1366, -282),]

final_remove<- lm(formula = target_death_rate ~ incidence_rate + poverty_percent + 
    pct_bach_deg25_over + pct_unemployed16_over + pct_white + 
    pct_other_race, data = cancer_df_remove)

#compare model with or without those three points
summary(final_model)
summary(final_remove) 

#all variables are still significant, and the parameter estimate is stable. Those three outliers are not influential points.
plot(final_remove)
```

We can see that the parameter estimates are almost the same with or without the three points. In this case, they are not influential points in final model. We decide to keep them in final model.

##Model Validation

###LOOCV
```{r}
glm.fit<-glm(target_death_rate ~ incidence_rate + poverty_percent + 
    pct_bach_deg25_over + pct_unemployed16_over + pct_white + 
    pct_other_race, data = cancer_data_df)

summary(glm.fit)

cv.err<-cv.glm(cancer_data_df,glm.fit)
cv.err$delta  

anova(final_model) %>% 
  tidy()
```

###Bootstrapping - residual sample

```{r}
pred = final_model$fitted.values
resid = final_model$residuals

#write a function for residual sampling
boots_trap = function(len) {
  rmse <- tibble()
  for (i in 1:len){
    sample_resid = sample(resid, 3047, replace = TRUE)
    t = cancer_data_df %>% 
      cbind(pred, sample_resid) %>% 
      mutate(new_target_death_rate = sample_resid + pred) %>% 
      lm(new_target_death_rate ~ incidence_rate + poverty_percent + 
    pct_bach_deg25_over + pct_unemployed16_over + pct_white + 
    pct_other_race, data = .) %>% 
      glance() %>% 
      dplyr::select(sigma)
  
    rmse = rmse %>% 
      rbind(t)
  }
  rmse
}

#run bootstrapping for 100, 1000, 10000 times
boots_100 = boots_trap(100)
boots_1000 = boots_trap(1000)
boots_10000 = boots_trap(10000)

boots_100 %>%
  summary() 
boots_1000 %>%
  summary() 
boots_10000 %>%
  summary() 
```